Cellular Automata and Computational Universality

### 2018-09-19 paper/paper.tex
Almost done

    \subsection{Fractran}
    \subsection{Langton's Ant}
\section{Applications}
    \subsection{Cryptography}
\section{Firing Squad Synchronisation Problem}
    \subsection{The Problem}
    Imagine a group of soldiers all part of a firing squad, with a general at
    one end. Firing squads would aim to have all of its soldiers fire at the
    same time so that no one soldier could be sure he had delivered the fatal
    wound to the condemned. The soldiers of the firing squad can be modelled as
    cells in a cellular automaton, and so the Firing Squad Synchronisation
    Problem, or FSSP, is concerned with ensuring that all states end up in a
    ``firing'' state at some point, at the same point in time. At each time
    step, a soldier may change what he is doing based on what him and his two
    immediate neighbours are doing (so the two soldiers on the ends only take
    into account their own current action and that of his single neighbour).
    Assume all soldiers start off in the same inactive state, also called the
    quiescent state. The question begins when the general initially gives the
    command to fire: assuming there is a fixed number of states, regardless of
    how many soldiers are in the line, how can the states be defined such that
    all soldiers are in the firing state at the same time?

\section{Majority Problem}

\bibliography{bibliography}

\end{document}

>>>>

    \subsection{Fractran}
    \subsection{Langton's Ant}
\section{Applications}
    \subsection{Cryptography}

\section{Firing Squad Synchronisation Problem}
    \subsection{The Problem}
    A firing squad consists of a group of soldiers, all of whom will fire a shot
    at the condemned party in order to end his or her life, at the command of a
    general. An important factor to this is that all the soldiers fire
    simultaneously; this means that no one soldier can be sure that they are the
    one to have administered the fatal shot, and can thus rest easy at night. If
    we model the line of soldiers as a one-dimensional cellular automaton, we
    want all of the cells in this automaton to be in a ``firing'' state
    simultaneously.  More specifically, at each time step each soldier may
    change what he is doing (i.e. change his state) based solely upon what he
    and his two immediate neighbours are doing. Note that the soldiers at either
    end of the line only consider the actions of himself and his sole neighbour.
    Furthermore, we assume that each soldier starts in the same inactive state
    (also known as the quiescent state), and that the general is located at one
    end of the line, and only he may give the initial command to fire. He
    himself must also fire his weapon at the to-be recipient of the discharge of
    such weapons. Thus the question asked by the Firing Squad Synchronisation
    Problem is this: assuming there is a fixed number of states, and regardless
    of how many soldiers form the firing squad, how can the states and
    transitions be defined such that all soldiers are all, at some moment in
    time, firing together?

    \subsection{The Solution}
    Since the signal can only travel at most one cell to the left/right between
    each time step, it is clear that a minimal-time solution can not take shorter
    than $2(n-1)$ time (the signal must travel all the way down the line and all
    the way back, on the way down skipping the general and on the way back
    bouncing straight off the end soldier). The first minimal-time solution to
    the problem was Abraham Waksman's, published in 1966 \cite{Waksman}. This
    solution uses the following six basic states:

    \begin{enumerate}
        \item $Q$ - the starting, quiescent state
        \item $T$ - the final, firing state
        \item $R$ - the trigger signal for the $B$ state:
            \begin{itemize}
                \item $R_0$ - propagates left
                \item $R_1$ - propagates right
            \end{itemize}
        \item $B$ - the state that generates the $P$ state
            \begin{itemize}
                \item $B_0$ - blocks the $R$ state
                \item $B_1$ - passes through the $R$
            \end{itemize}
        \item $P$ - the state generating the $A$ state (leads to $T$ if both
            neighbours in $P$ state)
            \begin{itemize}
                \item $P_0$ - generates $A_{0xx}$
                \item $P_1$ - generates $A_{1xx}$
            \end{itemize}
        \item $A$ - the propagating state
            \begin{itemize}
                \item $A_{000}, A_{001}, A_{010}, A_{011}, A_{100}, A_{101},
                    A_{110}, A_{111}$
                \item $A_{0xx}$ - generates the state $R$ with no delay
                \item $A_{1xx}$ - generates the state $R$ with one unit time
                    delay
                \item $A_{x00}, A_{x01}$ - propagates to the left
                \item $A_{x10}, A_{x11}$ - propagates to the right
            \end{itemize}
    \end{enumerate}

    There are also two extra placeholder states:
    \begin{itemize}
        \item $\phi$ - external state, no neighbour to this side
        \item $\gamma$ - neighbours note explicitly mentioned
    \end{itemize}

    The way Waksman's solution works can be summarized by the interactions
    between the states of his automaton which occur:
    \begin{itemize}
        \item Every $A$ signal can be assigned a parity, even or odd, based on
            whether it is at an even or odd cell. The parity is relative to the
            cell who originated the signal
        \item Every cell with an even $A$ signal generates a new $R$ signal that
            travels in the opposite direction of the cell's $A$ signal
        \item Each type of $B$ signal will switch to the other type upon
            intersection with an $R$ signal. One type allows the $R$ signal to
            continue propagating, the other does not
        \item When $A$ and $B$ signals intersect, they set up a new generator,
            or $P$ signal
    \end{itemize}

    Of course, time complexity is not the only metric by which to judge a
    solution's merit. The first solution to the problem, proposed by John
    McCarthy and Marvin Minsky, was relatively much more simple, involving
    propagating two waves down the line of soldiers, one moving at three times
    the speed of the other. The faster wave bounces off the edges and meet again
    in the center, creating two more waves for a total of four waves. This
    process repeats, until each cell is part of a wave simultaneously. At this
    point, the soldiers fire. This requires $3n$ units of time for $n$ soldiers,
    and requires 15 states.

    The current best solution to the problem was devised by Jacques Mazoyer
    \cite{FSSP_best}, and uses six states to solve the problem in time $2(n-1)$.

    In minimal-time solutions, the general sends signals $S_1, S_2, S_3, ..., S_i$ to
    the right (assuming he is the left most soldier) at speeds $1, \frac{1}{3},
    \frac{1}{7}, ..., \frac{1}{2^{i-1}-1}$. $S_1$ reflects and meets signal
    $S_i$ (for $i \ge 2$) at cell $\frac{n}{2^{i-1}}$. When $S_1$ reflects it also
    creates another general at the right end, and the process repeats. Fig.
    \ref{fig:mazoyer} shows Mazoyer's minimal-time and so far minimal-state
    solution to FSSP.

    \begin{figure}[h]
        \centering
        \includegraphics[width=4in]{mazoyer_fssp.png}
        \caption{Mazoyer's minimal solution to FSSP. Time increases from left to
        right.}
        \label{fig:mazoyer}
    \end{figure}

\section{Majority Problem}
    \subsection{The Problem}
    Suppose we have a one-dimensional cellular automaton, each of whose cells
    may be in one of two states. There are $i+j$ cells in total, $i$ of which
    are in the zero state, the other $j$ of which are in the one state. A
    correct solution to this problem must eventually set all cells to zero if $i
    > j$, or one if $i < j$. The desired final state of the automaton is
    unspecified if $i = j$.

    \subsection{The Solution}
    G\'acs, Kurdyumov, and Levin designed an automaton the solves the Majority
    Problem in most cases, but not all. In fact, with a random starting
    configuration, their solution is about 78\% successful in correctly
    determining the majority \cite{GKL_majority}. They approached it by
    classifying the quality of a particular cellular automaton by the proportion
    of the $2^{i+j}$ possible starting configurations in which it correctly
    identified the majority.

    An overview of their method is as follows: if a cell is 0, the next state is
    the majority value between itself, its neighbour immediately to its left,
    and its neighbour three spaces to the left. Similarly, if the cell is a 1,
    the value is taken as the majority of itself, its immediate right neighbour,
    and its neighbour three spaces to the right.

    Capcarrere, Sipper, and Tomassini observed \cite{CST_majority} the Majority
    Problem may be solved perfectly by altering the definition of having
    recognised the majority. For example, if Rule 184 is run on a finite
    universe with cyclic boundaries, each cell will eventually see two
    consecutive states of the majority value infinitely often, but will see two
    consecutive states of the minority value only finitely many times. Thus the
    Majority Problem cannot be solved perfectly if we require all cells to
    eventually stabilise to the majority state, but can if we relax this
    requirement and allow Rule 184 to run infinitely \cite{LB_majority}.

\bibliography{bibliography}
\end{document}
